Convective heat transport in stratified atmospheres at low and high Mach number

### 2017-03-17 tex/nusselt_paper.tex
pounded in a bunch of words about the 2D v 3D results

\begin{abstract}
Convection in astrophysical systems is stratified and
often occurs at high Rayleigh number (Ra) and low
Mach number (Ma).
Here we study stratified convection in the context of 
plane-parallel, polytropically stratified atmospheres. 
We hold the density stratification ($n_{\rho}$), aspect
ratio, and Prandtl 
number (Pr) constant while varying the superadiabaticity
and Ra.  We find that Ma is a strong function of the superadiabaticity
and a weak function of the Rayleigh number.
We also examine the behavior of the Nusselt number (Nu), 
which quantifies the efficiency of convective heat transport,
and the Reynolds number (Re), which quantifies how turbulent our
solutions are.  Both of these measures experience weaker scaling when
the evolved solution has an average Ma of one than they do in the
low Ma regime.
%As Ra increases and $\text{Ma} \rightarrow 1$, a scaling 
%of Nu $\propto$ Ra$^{0.45}$ is observed.  
%As Ra increases to a regime where Ma $\geq 1$,
%this scaling gives way to a weaker Nu $\propto$ Ra$^{0.19}$. 
%In the regime of Ma $\ll 1$, a consistent
%Nu $\propto$ Ra$^{0.31}$ is retrieved,  reminiscent of the 
%Nu $\propto$ Ra$^{2/7}$ seen in \RB convection.
\end{abstract}
\maketitle

@@

We find that the Mach number is a strong function of 
$\epsilon$ and a weak function of Ra.  
When Ma $< 1$, we observe a scaling law of
Ma(Ra$, \epsilon) \propto \epsilon^{1/2}$Ra$^{1/4}$, 
but this relation breaks down as the mean
Ma approaches 1 and the flows become supersonic (see Fig. \ref{fig:ma_v_eps}).  
Furthermore, the value of $\epsilon$ sets the size of the
evolved thermodynamic variables, such that
$T_1/T_0 \propto \epsilon$ and $\rho_1/\rho_0 \propto \epsilon$.  
These fluctuations are very small when $\epsilon$ and Ma are small,
but can be O(1) for values of $\epsilon$ near $m_{\text{ad}}$.

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/snapshots_fig.png}
\caption{Characteristic entropy fluctuations in evolved flows roughly
140$t_b$ after the start of simulations. 
The time- and horizontally-averaged profile is removed in all cases.  
(a) An $\epsilon = 10^{-4}$, low Ma flow. (b) An $\epsilon = 0.5$, flow,
in which shock systems can be seen in the upper atmosphere (e.g.
$(x, z) \approx (30, 18-19), (70, 15-19))$.
(c) A high Ra, $\epsilon = 0.5$ case.  Convective structures become
much smaller scale, and shock systems can be see near $(x, z) \approx (30, 16-19),
(65, 17-19)$.
\label{fig:entropy_snapshots} }
\end{figure}

@@

The variation of Nu with Ra is shown in 
Fig. \ref{fig:re_and_nu_v_ra}a.
At low to moderate Ra, 
Nu $\propto$ Ra$^{1/3}$ regardless of $\epsilon$,
reminiscent of scaling laws in classical \RB theory \cite{king&all2012}.
At large Ra, Nu $\propto$ Ra$^{1/5}$ 
at $\epsilon = 0.5$.
The scaling of Nu with Ra 
is unclear at low $\epsilon$ once Ra is sufficiently large
for shearing states to occur.  It is also important to note that
the value of Nu is heavily dependent upon the specific thermodynamic
structures of the solution, as double roll states will transport
heat more efficiently than single roll states, and slight changes in
Ra can result in a solution latching on to one solution or the other.

The Reynolds number (Re = $|\bm{u}|L_z/\nu$) and Peclet number
(Pe = Pr Re)
compare the importance of advection to diffusion in the evolved
convective state.  For Pr = 1, such as in this work, Pe = Re.  
Our choice of $\{\nu,\chi\}\propto \rho_0^{-1}$ drastically changes
the value of Re between the top and bottom of the atmosphere.  We report values of
Re at the midplane ($z=L_z/2$) of the atmosphere in
Fig. \ref{fig:re_and_nu_v_ra}b.
At low Ra, Re $\propto$ Ra$^{3/4}$.  At high Ra and $\epsilon$, 
where the average Ma $\approx 1$, 
this scaling gives way to a Re $\propto$ Ra$^{1/2}$,
where all changes in Re are due to the lowering of the diffusivity
through raising Ra.
For large values of Ra but at low $\epsilon$, there appears to be a scaling
of Re $\propto$ Ra$^{2/3}$, but once again the shearing states have made
these points difficult to measure.

@@

As the thermodynamic variables converge to their steady state values, 
the density profile evolves while remaining in hydrostatic equilibrium 
to zeroth order.  In Fig. \ref{fig:nrho_v_ra} we show the number of density 
scale heights present in the evolved solution using two measures.  
We find that once the average Ma of the domain
becomes approximately one, large density inversions begin to 
form in the boundary layers, as was reported by \cite{brandenburg&all2005}.  
The agreement of Nu across $\epsilon$ (Fig. \ref{fig:re_and_nu_v_ra}a), 
particularly at low Ra in which all four of our cases collapse onto a single
power law, is striking in light of the vastly different evolved stratifications
felt by the flows.

\section{Discussion}
\refstepcounter{section}
\label{sec:discussion}
In this letter we have studied fundamental heat transport by 
stratified convection in simplified 2D and 3D polytropic atmospheres.
We argue that these atmospheres are the natural extension
of the RB problem to stratified systems, 
and are an ideal laboratory for understanding the basic 
properties of stratified convection. 
We see little difference between
the properties of our 2D and 3D results, which aligns with expectations
of Boussinesq theory and numerical simulation results at 
values of Pr $\geq$ 1 \cite{ahlers&all2009}.

>>>>

\begin{abstract}
Convection in astrophysical systems is stratified and
often occurs at high Rayleigh number (Ra) and low
Mach number (Ma).
Here we study stratified convection in the context of 
plane-parallel, polytropically stratified atmospheres. 
We hold the density stratification ($n_{\rho}$), aspect
ratio, and Prandtl 
number (Pr) constant while varying the superadiabaticity
and Ra.  We find that Ma is a strong function of the superadiabaticity
in both 2D and 3D simulations, and that it is a 
weak function of Ra in 2D.
In 2D simulations, the scaling of the Nusselt Number (Nu) and the Reynolds number (Re)
with Ra weaken
once the mean Ma of the solution is unity, but in 3D
Re and Nu seem to have consistent scaling across Ra.
We further find that when the superadiabaticity and Ma are large,
density inversions form in the thermal boundary layers.
%As Ra increases and $\text{Ma} \rightarrow 1$, a scaling 
%of Nu $\propto$ Ra$^{0.45}$ is observed.  
%As Ra increases to a regime where Ma $\geq 1$,
%this scaling gives way to a weaker Nu $\propto$ Ra$^{0.19}$. 
%In the regime of Ma $\ll 1$, a consistent
%Nu $\propto$ Ra$^{0.31}$ is retrieved,  reminiscent of the 
%Nu $\propto$ Ra$^{2/7}$ seen in \RB convection.
\end{abstract}
\maketitle

@@

We find that the Mach number is a strong function of 
$\epsilon$ and a weak function of Ra.  
In our 2D runs, when Ma $< 1$ we observe a scaling law of
Ma(Ra$, \epsilon) \propto \epsilon^{1/2}$Ra$^{1/4}$, 
but this relation breaks down as the mean
Ma approaches 1 and the flows become supersonic (see Fig. \ref{fig:ma_v_eps}).  
In 3D, Ma appears to be a function of $\epsilon$ alone, with
Ma $\propto \epsilon^{1/2}$.
Furthermore, the value of $\epsilon$ sets the size of the
evolved thermodynamic variables, such that
$T_1/T_0 \propto \epsilon$ and $\rho_1/\rho_0 \propto \epsilon$.  
These fluctuations are very small when $\epsilon$ and Ma are small,
but can be O(1) for values of $\epsilon$ near $m_{\text{ad}}$.

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/snapshots_fig.png}
\caption{Characteristic entropy fluctuations in evolved 2D flows roughly
140$t_b$ after the start of simulations. 
The time- and horizontally-averaged profile is removed in all cases.  
(a) An $\epsilon = 10^{-4}$, low Ma flow. (b) An $\epsilon = 0.5$, flow,
in which shock systems can be seen in the upper atmosphere (e.g.
$(x, z) \approx (30, 18-19), (70, 15-19))$.
(c) A high Ra, $\epsilon = 0.5$ case.  Convective structures become
much smaller scale, and shock systems can be see near $(x, z) \approx (30, 16-19),
(65, 17-19)$.
\label{fig:entropy_snapshots} }
\end{figure}

@@

The variation of Nu with Ra is shown in 
Fig. \ref{fig:re_and_nu_v_ra}a.
In 2D and at low to moderate Ra, 
Nu $\propto$ Ra$^{1/3}$ regardless of $\epsilon$,
reminiscent of scaling laws in classical \RB theory \cite{king&all2012}.
At large Ra, Nu $\propto$ Ra$^{1/5}$ 
at $\epsilon = 0.5$.
The scaling of Nu with Ra 
is unclear at low $\epsilon$ once Ra is sufficiently large
for shearing states to occur.  It is also important to note that
the value of Nu is heavily dependent upon the specific thermodynamic
structures of the solution, as double roll states will transport
heat more efficiently than single roll states, and slight changes in
Ra can result in a solution latching on to one solution or the other.
In 3D, it appears that Nu $\propto$ Ra$^{2/7}$, a classic scaling law
seen in RB studies \cite{johnston&doering2009}.

The Reynolds number (Re = $|\bm{u}|L_z/\nu$) and Peclet number
(Pe = Pr Re)
compare the importance of advection to diffusion in the evolved
convective state.  For Pr = 1, such as in this work, Pe = Re.  
Our choice of $\{\nu,\chi\}\propto \rho_0^{-1}$ drastically changes
the value of Re between the top and bottom of the atmosphere.  We report values of
Re at the midplane ($z=L_z/2$) of the atmosphere in
Fig. \ref{fig:re_and_nu_v_ra}b.
In 2D and at low Ra, Re $\propto$ Ra$^{3/4}$.  At high Ra and $\epsilon$, 
where the average Ma $\approx 1$, 
this scaling gives way to a Re $\propto$ Ra$^{1/2}$,
where all changes in Re are due to the lowering of the diffusivity
through raising Ra.
For large values of Ra but at low $\epsilon$, there appears to be a scaling
of Re $\propto$ Ra$^{2/3}$, but once again the shearing states have made
these points difficult to measure.  In 3D, the scaling of Re appears to
consistently be Re $\propto$ Ra$^{1/2}$, consistent with the high-$\epsilon$,
high Ra results seen in 2D and the lack of scaling of Ma with Ra in Fig. 
\ref{fig:ma_v_eps}b.

@@

As the thermodynamic variables converge to their steady state values, 
the density profile evolves while remaining in hydrostatic equilibrium 
to zeroth order.  In Fig. \ref{fig:nrho_v_ra} we show the number of density 
scale heights present in the evolved solution using two measures.  
We find that in 2D, once the average Ma of the domain
becomes approximately one, large density inversions begin to 
form in the boundary layers, as was reported by \cite{brandenburg&all2005}.  
The agreement of Nu across $\epsilon$ (Fig. \ref{fig:re_and_nu_v_ra}a), 
particularly at low Ra in which all four of our cases collapse onto a single
power law, is striking in light of the vastly different evolved stratifications
felt by the flows.  In 3D, despite the Ma not scaling with Ra, we see similar density
inversions at high $\epsilon$ once Ra is sufficiently large.

\section{Discussion}
\refstepcounter{section}
\label{sec:discussion}
In this letter we have studied fundamental heat transport by 
stratified convection in simplified 2D and 3D polytropic atmospheres.
We argue that these atmospheres are the natural extension
of the RB problem to stratified systems, 
and are an ideal laboratory for understanding the basic 
properties of stratified convection. 
The primary difference we see between our 2D and 3D results is that the Ma
is not a function of Ra in 3D, and this has far-reaching consequences.  Otherwise,
we see only slightly different scaling in Nu(Ra) in 2D and 3D, 
which aligns with expectations
of Boussinesq theory and numerical simulation results at 
values of Pr $\geq$ 1 \cite{ahlers&all2009}.
